Observation of the orientation of snow hydrometeors at sheltered and unsheltered sites

### 2024-12-16 Review_R1_answers_202412.tex

    We agree with the reviewer that the distribution of wind components and turbulence at the relevant scales is probably more important than simple wind speed that we could investigate in this work (as all the MASC measurements in the database have co-located wind speed information available).

    We carefully revised the manuscript to make sure not to overstate the role of wind speed alone and also to clarify when we cannot separate between the "natural" contribution of wind on the distribution of orientations and the "artificial" contribution due to the turbulence generated by strong wind interacting with the hollow measurement system of the MASC. A few changes are highlighted in the manuscript. We may cite for example in Sec.~3:

    \blue{Atmospheric conditions are expected to play a role in shaping orientation distributions: atmospheric turbulence is probably a key driver, as well as horizontal winds interacting with the instrument frame~\cite{Fitch_AMT_2021}. Turbulence has been shown to heavily impact the falling of particles in fluids in general~\cite{Petersen_JFM_2019,Brandt_ARFM_2022} and of frozen hydrometeors in particular~\cite{Nemes_JFM_2017,Li_JFM_2021}.}

     Later in the same section:

    \blue{This interpretation cannot unfortunately tell us any relevant information about the role of environmental turbulence, but only the likely effect of local turbulence introduced by the instrument.  Further research is therefore necessary to clarify these dynamics, in particular the role of natural turbulence in broadening the orientation distribution.}

    And in the Conclusions:

    \blue{It is clear that further research, with different setups is needed to separate the contributions of these effects and estimate the \textit{natural} effect of wind and turbulence on the distribution of orientation values. For this purpose, instruments that do not create local perturbations (like PIP) could be used, ideally co-located with a MASC. }
\end{enumerate}

@@

    \noindent
    We agree with the reviewer, and we implemented their suggestion. \\

    \item \textit{Introduction: Honestly, comparing it with the other sections, I find the introduction a bit weak. Considering the wider audience addressed by GRL, I think the importance of ice particle orientation should be emphasized and explained more. It is certainly true that most radar polarimetric quantities are sensitive to orientation (L. 50-52). But this is by far not all in my opinion. Particle orientation also impacts the scattering signal in passive microwave (MW) observations (both from space and from ground). Many space-borne MW sensors are also polarization-sensitive. Considering the enormous efforts undertaken for example by ECMWF and other services to assimilate those all-sky MW observations, this aspect should be mentioned and cited. Another aspect is that an ice particle falling with a preferred orientation will have a very different sedimentation velocity. So characterizing the particle orientation is quite key to simulate its terminal velocity (which is quite an important parameter in any cloud model). Ice particle orientation is also relevant for determining cloud radiative properties such as optical depth or albedo.}

    We thank the reviewer for the suggestion. While we left the second part of the Introduction a bit more "specialized" (targeting an audience interested in snow imagers and their data processing), we realized that we did not provide a satasfactory broad overview, neglecting all the implications listed by the reviewer. We rephrased therefore the first part of the Introduction, removed a few technical paragraphs and included most of the suggestions of the reviewer. Some new paragraphs now read:

    \blue{The preferential orientation of falling hydrometeors is exploited by dual-polarization weather radars to retrieve information about the microphysical processes occurring in precipitation and to identify which type of hydrometeors are present and where. This is relevant for the interpretation of ground-based polarimetric radar data as well as for exploiting satellite-based observations and their assimilation, as done for example by the European Centre for Medium-Range Weather Forecasts ECMWF~\cite{Eyre_QJRMS_2022}. For example, recent studies showed that considering hydrometeor orientation in the assimilation of high-frequency radiometry may improve forecast in areas of cloud and precipitation~\cite{Barlakas_AMT_2021} and produce significant improvements in path-integrated quantities such as the Ice Water Path~\cite{Kaur_RS_2022}. }

@@

    \blue{The orientation and falling regime also affects the settling velocity of ice phase hydrometers, a key parameter linking microphysical properties to other tangible quantities as the precipitation intensity, and that remains to be better understood and documented~\cite{Heymsfield_JAS_2004}.}\\

    \item \textit{L. 46: It would be good to add a reference for this statement. I question a bit whether this general statement is actually true. I completely agree for plate-like crystals or large aggregates. But what is for example about bullet rosettes in high cirrus clouds? Or columns/needles? Do we have solid knowledge about their falling orientation? Also the references you provide a few sentences later (\cite{Noel_JAMC_2005,Tinklenberg_JFM_2023}) focus only on planar crystals.}

    We believe the reviewer is right. We were biased tp think mostly in terms of larger snow particles (snow aggregates, graupel) or planar crystals, so our formulation was overstating the current knowledge. We rephrased fully this sentence as:\\

    \blue{Ice-phase hydrometeors are typically assumed to fall with a preferred orientation. Falling snow particles populating frozen clouds are  considered to align preferentially in a way that maximizes the area projected in the vertical direction, as presented in studies conducting remote sensing retrievals and laboratory experiments.  Such experiments are documented for planar crystals~\cite{Noel_JAMC_2005,Matrosov_JAS_2005,Tinklenberg_JFM_2024} as well as for more complex and larger snowflakes and rimed hydrometeors~\cite{Kennedy_JAMC_2011, Ryzhkov_JAMC_2011, Koebschall_EF_2023}, while less is documented for other particle types such as needles, columns or rosettes. }\\

### 2024-12-16 Grazioli_JGRA_2024_R1.tex
Updates from Overleaf

\section{Introduction}
Ice-phase hydrometeors are typically assumed to fall with a preferred orientation. Falling snow particles populating frozen clouds are generally considered to align preferentially in a way that maximizes the area projected in the vertical direction, as presented in studies conducting remote sensing retrievals and laboratory experiments. Such experiments are documented for planar crystals~\cite{Noel_JAMC_2005,Matrosov_JAS_2005,Tinklenberg_JFM_2024} as well as for more complex and larger snowflakes and rimed hydrometeors~\cite{Kennedy_JAMC_2011, Ryzhkov_JAMC_2011, Koebschall_EF_2023}, while less is documented for other particle types such as needles, columns or rosettes.

The preferential orientation of falling hydrometeors is exploited by dual-polarization weather radars to retrieve information about the microphysical processes occurring in precipitation and to identify which type of hydrometeors are present and where. This is relevant for the interpretation of ground-based polarimetric radar data as well as for exploiting satellite-based observations and their (all-sky) assimilation, as done for example by the European Centre for Medium-Range Weather Forecasts ECMWF~\cite<e.g.,>{Eyre_QJRMS_2022}. For example, recent studies showed that considering hydrometeor orientation in the assimilation of high-frequency radiometry may improve forecast in areas of cloud and precipitation~\cite{Barlakas_AMT_2021} and produce significant improvements in path-integrated quantities such as the Ice Water Path~\cite{Kaur_RS_2022}.

@@

Considering a population of hydrometeors rather than individual particles, as is often the case for practical applications, the distribution of orientations is often modeled with relatively narrow variability around 0$^\circ$ values (i.e., horizontally oriented). Looking at the overview in Table~\ref{table:orientations}, one can observe that modeling studies mostly assume a Gaussian distribution, with standard deviations varying between 3$^\circ$ \cite{Matrosov_JAM_2001} and 60$^\circ$ \cite{Putnam_MWR_2017} depending on the type of hydrometeors. Observational studies conducted in the field \cite{Garrett_GRL_2015} confirm that there are differences due to the type of hydrometeor and environmental conditions (wind and turbulence). They also report non-zero mean and median orientation values, which is contradictory with what was stated previously. As we will discuss later in this study, this discrepancy can be the result of the way multi-view instrumental observations are combined to retrieve an overall estimate of orientation, at least for more complex three-dimensional snow particles. A recent laboratory study using snow crystals replicas falling in an alcoholic solution~\cite{Stout_ACP_2024}, however, shows that in quiescent conditions individual planar crystals may have a spiralling fall regime with preferential orientations that depart from 0$^\circ$. A study with similar methods, but focused on larger snow aggregates, instead showed that horizontal orientation is preferred for these hydrometeors~\cite{Koebschall_EF_2023}. It is to be remarked that those and other studies, in considering objects falling in liquids, use a solid-to-fluid density ratio which is orders lower than in atmospheric precipitation. This may have far-reaching consequences in the settling dynamics~\cite{Tinklenberg_JFM_2023,Tinklenberg_JFM_2024}.

>>>>

\section{Introduction}
Ice-phase hydrometeors are typically assumed to fall with a preferred orientation. Falling snow particles populating frozen clouds are considered to align preferentially in a way that maximizes the area projected in the vertical direction, as presented in studies conducting remote sensing retrievals and laboratory experiments. Such experiments are documented for planar crystals~\cite{Noel_JAMC_2005,Matrosov_JAS_2005,Tinklenberg_JFM_2024} as well as for more complex and larger snowflakes and rimed hydrometeors~\cite{Kennedy_JAMC_2011, Ryzhkov_JAMC_2011, Koebschall_EF_2023}, while less is documented for other particle types such as needles, columns or rosettes.

The preferential orientation of falling hydrometeors is exploited by dual-polarization weather radars to retrieve information about the microphysical processes occurring in precipitation and to identify which type of hydrometeors are present and where. This is relevant for the interpretation of ground-based polarimetric radar data as well as for exploiting satellite-based observations and their assimilation, as done for example by the European Centre for Medium-Range Weather Forecasts ECMWF~\cite<e.g.,>{Eyre_QJRMS_2022}. For example, recent studies showed that considering hydrometeor orientation in the assimilation of high-frequency radiometry may improve forecast in areas of cloud and precipitation~\cite{Barlakas_AMT_2021} and produce significant improvements in path-integrated quantities such as the Ice Water Path~\cite{Kaur_RS_2022}.

@@

Considering a population of hydrometeors rather than individual particles, the distribution of orientations is often modeled with relatively narrow variability around 0$^\circ$ values (i.e., horizontally oriented). Looking at the overview in Table~\ref{table:orientations}, one can observe that modeling studies mostly assume a Gaussian distribution, with standard deviations varying between 3$^\circ$ \cite{Matrosov_JAM_2001} and 60$^\circ$ \cite{Putnam_MWR_2017} depending on the type of hydrometeors. Observational studies conducted in the field \cite{Garrett_GRL_2015} confirm that there are differences due to the type of hydrometeor and environmental conditions (wind and turbulence). They also report non-zero mean and median orientation values, which is contradictory with what was stated previously. As we will discuss later in this study, this discrepancy can be the result of the way multi-view instrumental observations are combined to retrieve an overall estimate of orientation, at least for more complex three-dimensional snow particles. A recent laboratory study using snow crystals replicas falling in an alcoholic solution~\cite{Stout_ACP_2024}, however, shows that in quiescent conditions individual planar crystals may have a spiraling fall regime with preferential orientations that depart from 0$^\circ$. A study with similar methods, but focused on larger snow aggregates, instead showed that horizontal orientation is preferred for these hydrometeors~\cite{Koebschall_EF_2023}. It is to be remarked that those and other studies, in considering objects falling in liquids, use a solid-to-fluid density ratio which is orders lower than in atmospheric precipitation. This may have far-reaching consequences in the settling dynamics~\cite{Tinklenberg_JFM_2023,Tinklenberg_JFM_2024}.
